---
title: "intro-geoda-r"
author: "lixun910"
date: "6/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install libgeoda for R

For testing stage, the easiest way to install the development version of libgeoda is using `devtools`. If you don’t have devtools installed, use the following line:
```r
install.packages("devtools")
```

Then, you can call `install_githu()` function to install `libgeoda` from the source code:

```r
library(devtools)
install_github('lixun910/libgeoda-r')
```

If everything installed without error, you should be able to load `libgeoda`:

```{r}
library(libgeoda)
```


## Load Spatial Data

There is an ESRI Shapefile `natregimes.shp` comes with the package:

```{r}
nat_path <- system.file("extdata", "natregimes.shp", package = "libgeoda")
nat_path
```


Use the path above (or the path to your own dataset), we can create a GeoDa instance, and the GeoDa instance is the main entry point of libgeoda. 


```{r}
nat <- GeoDa(nat_path)
```

There are some functions provided to check the meta data of the loaded dataset:

```{r}
num_obs <- nat$GetNumObs()
num_cols <- nat$GetNumCols()

field_types <- nat$GetFieldTypes()
field_names <- nat$GetFieldNames()
field_names
```


## --Create GeoDa instance from SF (simple feature) instance--

## Table Data

Using function `GetNumericCol(string col_name)` to get numeric data (vector type) from column "HR60":

```{r}
data <- nat$GetNumericCol("HR60")
data[1:20]
```

## Spatial Weights

Using function `CreateQueenWeights(string poly_id="", int order=1, bool include_lower_order=false)` to create a Queen contiguity weights:
```{r}
queen_w <- nat$CreateQueenWeights()
cat("weight_type: ", queen_w$weight_type, "\nis_symmetric: ", queen_w$is_symmetric, "\nsparsity:",queen_w$sparsity, "\ndensity:",queen_w$density, "\nmin_nbrs:",queen_w$min_nbrs, "\nmax_nbrs:",queen_w$max_nbrs, "\nmean_nbrs:",queen_w$mean_nbrs, "\nmedian_nbrs:",queen_w$median_nbrs, "\n")
```

You can save the weights file by calling function `SaveToFile(string file_path)`:

## Spatial Data Analysis

### Local Spatial Autocorrelation

Using the created Queen weights and the data "HR60", we can call function `LISA(GeoDaWeight w, vector data)` to compute the local spatial autocorrelation of variable "HR60". We can access the LISA results by calling the "getter" methods from the returned LISA object


```{r}
lisa <- nat$LISA(queen_w, data)
print("Spatial Lag values:")
lags <- lisa$GetLagValues()
lags[1:10]
lms <- lisa$GetLocalMoranValues()
lms[1:10]
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
cats <- lisa$GetClusterIndicators()
cats[1:20]
```

You can easily re-run the LISA computation by calling its `Run()` function. For example, re-run the above LISA exampe using 9999 permutations:

```{r}
lisa$SetNumPermutations(9999)
lisa$Run()
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
```

Since libgeoda is using GeoDa's C++ code, by default, libgeoda uses multi-threading to accelerate the computation of Local Moran. One can also specify how many threads to run the computation:

```{r}
lisa$SetNumThreads(4)
lisa$Run()
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
```


## Clustering

### SKATER

Using the function `SKATER(int k, GeoDaWeight w, Vector column_names, String distance_method='euclidean')` to run a spatially constrained clustering on current dataset. 

```{r}
skater <- nat$SKATER(10, queen_w, c("HR60", "PO60"))
skater
```






