---
title: "ESDA_using_rgeoda_sf"
author: "Xun Li"
date: "1/13/2021"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{esda_using_rgeoda_sf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1 Exploratory Spatial Data Analysis basics

## 2 Exploratory Spatial Data Analysis with sf


In this section, we will work with sf library to do exploratory spatial data analysis (ESDA).

### 2.1 Start from `sf` package

The sf package has been popular tool to handle geospatial data. It is a good substitue of sp package which will be deprecated soon.

For example, to load a geospatial data (e.g. a ESRI Shapefile) using sf:

```{r, class.source='rCode'}
library(rgeoda)
guerry_path <- system.file("extdata", "Guerry.shp", package = "rgeoda")

library(sf)
guerry_sf <- st_read(guerry_path)
```
 
You can simply call plot() function to render the first 9 chorepleth maps using the frist 9 variables in the dataset:

```{r, class.source='rCode',fig.width = 6, fig.height=6}
plot(guerry_sf)
```

### 2.2 Create rgeoda object from `sf`


`sf` package makes it easy to work with geospatial data. It also provides functions to do basic spatial data analysis. `rgeoda` provides helper function `sf_to_geoda` to create a GeoDa instance from a `sf` object. Users can then call GeoDa functions to do further spatial data analysis.

Create a geoda object from the sf object guerry_sp you just created.


```{r, class.source='rCode'}
guerry_gda <- sf_to_geoda(guerry_sf)
```

rgeoda uses wkb, which is a binary representation of geometries, to exchange data between sf and libgeoda in memory.

By default, sf_to_geoda() only using geometries to create a geoda object, which is fast and user can reuse the data.frame from the sf object to do analysis. This function has an optional parameter with_table, to allow user to use both geometries and data.frame to create a geoda object by calling, for example: sp_to_geoda(guerry_sp, with_table=TRUE)

### 2.3 ESDA with rgeoda

Now, with the geoda object guerry, you can call rgeoda's spatial analysis functions. For example, to examine the local Moran statistics of variable "crm_prs" (Population per Crime against persons):


```{r, class.source='rCode'}
queen_w <- queen_weights(guerry_gda)

guerry_df <- as.data.frame(guerry_sf)
crm_prs <- guerry_df['Crm_prs'][,1]
```

#### 2.3.1 Local Moran Map

```{r}
lisa <- local_moran(queen_w, crm_prs)
```

Now, with the LISA results, we can do exploratory spatial data analysis by generating a LISA cluster map:

```{r, class.source='rCode', fig.width = 6, fig.height=6}
lisa_colors <- lisa_colors(lisa)
lisa_labels <- lisa_labels(lisa)
lisa_clusters <- lisa_clusters(lisa)

plot(st_geometry(guerry_sf), 
     col=sapply(lisa_clusters, function(x){return(lisa_colors[[x+1]])}), 
     border = "#333333", lwd=0.2)
title(main = "Local Moran Map of Crm_prs")
legend('bottomleft', legend = lisa_labels, fill = lisa_colors, border = "#eeeeee")
```

From the above code, you can see that we still use `sf` object to do plotting. The values of cluster indicators from `rgeoda`'s `LISA` object are used to make the LISA map.

You can easily append the lisa results to original `sf` object by manipulating the data.frame object inside the `sf` object.
```{r}
guerry_sf$moran_cluster <- lisa_clusters

```

If you check the values of the cluster indicators, you will see they are integer numbers 0 (not significant), 1 (high-high cluster), 2 (low-low cluster), 3 (low-high cluster), 4 (high-low cluster), 5 (neighborless/island), 6 (undefined), which are excatly the same with GeoDa software when you save LISA results to a table:


```{r}
lisa_clusters

```

To create a siginificant map that is associated with the Local Moran map:

```{r, fig.width = 6, fig.height=6}
lisa_p <- lisa_pvalues(lisa)
p_labels <- c("Not significant", "p <= 0.05", "p <= 0.01", "p <= 0.001")
p_colors <- c("#eeeeee", "#84f576", "#53c53c", "#348124")
plot(st_geometry(guerry_sf), 
     col=sapply(lisa_p, function(x){
       if (x <= 0.001) return(p_colors[4])
       else if (x <= 0.01) return(p_colors[3])
       else if (x <= 0.05) return (p_colors[2])
       else return(p_colors[1])
       }), 
     border = "#333333", lwd=0.2)
title(main = "Local Moran Map of Crm_prs")
legend('bottomleft', legend = p_labels, fill = p_colors, border = "#eeeeee")
```



