---
title: "ESDA with rGeoDa and sf"
author: "lixun910"
date: "7/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Previous R note shows you: 

how to do ESDA with rGeoDa and sp by computing the local spatial statistics (LISA) of variable `hr60` and creating a LISA cluster map.

Another modern and popular routine for spatial data analysis with R is to use `sf` library to process geospatial dataset, analyze spatial data and plot the results on maps.

`rgeoda` also provides utility functions to create a rgeoda instance from `sf` object.

In this note, I will demonstrate how to work with `sf` and `rgeoda` packages to do exploratory spatial data analysis.

## 1. Start from `sf` package

The sf package has been popular tool to handle geospatial data. It is a good substitue of `sp` package which will be deprecated soon. 

One can use sf to load geospatial data (e.g. a ESRI Shapefile) directly:

```{r}
library(rgeoda)
guerry_path <- system.file("extdata", "Guerry.shp", package = "rgeoda")
```

```{r}
library(sf)
guerry_sf <- st_read(guerry_path)
```
 
You can simply call plot() function to render the first 9 chorepleth maps using the frist 9 variables in the dataset:

```{r}
plot(guerry_sf)
```

## 2. Create rgeoda object from `sp`

`sf` package makes it easy to work with geospatial data. It also provides functions to do basic spatial data analysis. `rgeoda` provides helper function `sf_to_geoda` to create a GeoDa instance from a `sf` object. Users can then call GeoDa functions to do further spatial data analysis.

```{r}
guerry_gda <- sf_to_geoda(guerry_sf)
```

rgeoda uses wkb, which is a binary representation of geometries, to exchange data between `sf` and `libgeoda` in memory. 

## ESDA with rgeoda

Now, with the rgeoda object `guerry_gda`, you can call GeoDa's spatial analysis functions. For example, to examine the local Moran of variable "crm_prs" (Population per Crime against persons):

```{r}
queen_w <- guerry_gda$CreateContiguityWeights()
crm_lisa <- guerry_gda$LISA(queen_w, as.numeric(paste(guerry_sp$Crm_prs)))
crm_lisa_clusters <- crm_lisa$GetClusterIndicators()
```

Now, with the LISA results, we can do exploratory spatial data analysis by generating a LISA cluster map:

```{r}
plot(guerry_sp, 
     col=sapply(crm_lisa_clusters, function(x){return(lisa_colors[[x+1]])}), 
     border = "#333333", lwd=0.2)
title(main = "Local Moran Map")
legend('bottomright', 
       legend = lisa_labels, 
       fill = lisa_colors, border = "#eeeeee")
```
From the above code, you can see that we still use `sf` object to do plotting. The values of cluster indicators from `rgeoda`'s `LISA` object are used to make the LISA map.

You can easily append the lisa results to original `sf` object by manipulating the data.frame object inside the `sf` object.

If you check the values of the cluster indicators, you will see they are integer numbers 0 (not significant), 1 (high-high cluster), 2 (low-low cluster), 3 (low-high cluster), 4 (high-low cluster), 5 (neighborless/island), 6 (undefined), which are excatly the same with GeoDa software when you save LISA results to a table:

```{r}
crm_lisa_clusters
```

We also specify what color is associated with which cluster value to replace the `lisa_colors` and `lisa_labels`, which are defined in `sf_geoda.R`:
```
lisa_colors <- c("#eeeeee","#FF0000","#0000FF","#a7adf9", "#f4ada8", "#999999")
```
We then create an array of hex color, each observation is set to a hex color according to its cluster indicator value.
```{r}
sapply(crm_lisa_clusters, function(x){return(lisa_colors[[x+1]])})
```

