---
title: "Introduction to rGeoDa"
author: "lixun910"
date: "6/6/2019"
output: html_document
---

```{css}
.rCode {
background-color: #D7ECE8;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install rgeoda for R

For testing stage, the easiest way to install the development version of rgeoda is using `devtools`. If you don’t have devtools installed, use the following line:
```
install.packages("devtools")
```

Then, you can call `install_github()` function to install `rgeoda` from the source code:

```r
library(devtools)
install_github('lixun910/rgeoda')
```

If everything installed without error, you should be able to load `rgeoda`:

```{r, class.source='rCode'}
library(rgeoda)
```


## Load Geospatial Data

In this note, we will use an ESRI Shapefile `natregimes.shp` comes with the package:

```{r, class.source='rCode'}
nat_path <- system.file("extdata", "natregimes.shp", package = "rgeoda")
```


Use the path above (or the path to your own dataset), we can create a GeoDa instance, and the GeoDa instance is the main entry point of rgeoda. 


```{r, class.source='rCode'}
gda <- GeoDa(nat_path)
```

rGeoDa provides some functions to check the meta data of the loaded dataset:

```{r, class.source='rCode'}
num_obs <- gda$GetNumObs()
num_cols <- gda$GetNumCols()
field_types <- gda$GetFieldTypes()
field_names <- gda$GetFieldNames()
```


## Access Table Data

One can call function `GetNumericCol(string col_name)` to get numeric data (vector type) from GeoDa instance. For example, to get data of column "HR60":

```{r, class.source='rCode'}
hr60 <- gda$GetNumericCol("HR60")
```

## Spatial Weights

One can call function `CreateContiguityWeights(string poly_id="", bool is_queen=true, int order=1, bool include_lower_order=false)` to create a Queen or Rook contiguity weights:
```{r, class.source='rCode'}
queen_w <- gda$CreateContiguityWeights(is_queen=TRUE)
```

To access the properties of the created weights object, one can just access the attributes of the returned weights object `queen_w`:

```{r, class.source='rCode'}
cat("weight_type: ", queen_w$weight_type, 
    "\nis_symmetric: ", queen_w$is_symmetric, 
    "\nsparsity:",queen_w$sparsity,
    "\ndensity:",queen_w$density, 
    "\nmin_nbrs:",queen_w$min_nbrs, 
    "\nmax_nbrs:",queen_w$max_nbrs,
    "\nmean_nbrs:",queen_w$mean_nbrs, 
    "\nmedian_nbrs:",queen_w$median_nbrs, "\n")
```

To save the weights file to local disk, one can call the function of weights object `SaveToFile(string file_path)`

## Spatial Data Analysis

### Local Spatial Autocorrelation

Using the created Queen weights `queen_w` and the data `hr60`, we can call function `LISA(GeoDaWeight w, vector data)` to compute the local spatial autocorrelation of variable "HR60". 


```{r, class.source='rCode'}
lisa <- gda$LISA(queen_w, hr60)
```

We can access the LISA results by calling the "getter" methods from the returned LISA object:
```{r, class.source='rCode'}
lags <- lisa$GetLagValues()
lags[1:20]
```

Get local moran values:
```{r, class.source='rCode'}
lms <- lisa$GetLocalMoranValues()
lms[1:20]
```

Get pseudo-p values:
```{r, class.source='rCode'}
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
```

Get LISA category values:

```{r, class.source='rCode'}
cats <- lisa$GetClusterIndicators()
cats[1:20]
```

You can easily re-run the LISA computation by calling its `Run()` function. For example, re-run the above LISA exampe using 9999 permutations:

```{r, class.source='rCode'}
lisa$SetNumPermutations(9999)
lisa$Run()
```

Display the p-values after 9999 permutations:
```{r, class.source='rCode'}
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
```

Since rgeoda is using GeoDa's C++ code, by default, rgeoda uses multi-threading to accelerate the computation of Local Moran. One can also specify how many threads to run the computation:

```{r, class.source='rCode'}
lisa$SetNumThreads(4)
lisa$Run()
```


Display the p-values after re-run LISA with 4 threads:
```{r, class.source='rCode'}
pvals <- lisa$GetLocalSignificanceValues()
pvals[1:20]
```


## Clustering

### SKATER

Using the function `SKATER(int k, GeoDaWeight w, Vector column_names, String distance_method='euclidean')` to run a spatially constrained clustering on current dataset. 

```{r, class.source='rCode'}
skater <- gda$SKATER(10, queen_w, c("HR60", "PO60"))
skater
```

## Visualization using sp

```{r, class.source='rCode'}
spo <- geoda_to_sp(gda)
colors <- sapply(cats, function(x){return(lisa_colors[[x+1]])})
plot(spo, col=colors, border = "#333333", lwd=0.2)
title(main = "LISA Map")
legend('bottomright', legend = lisa_labels, 
       fill = lisa_colors, border = "#eeeeee")
```






