---
title: "rGeoDa and sp (dev notes 1)"
author: "lixun910"
date: "6/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`libgeoda for R` (or rgeoda) is a R package that wraps all core functions of spatial data analysis in GeoDa. Unlike the desktop software GeoDa, libgeoda is a non-UI and feature focused C++ library that is designed for programmers to do spatial data analysis using their favoriate programming languages (R, Python, Java etc.). It also aims to be easily integratd with other libraries, softwares or systems on different platforms. 

Previous "intro" page shows you: 

1. how to install rgeoda, 
2. how to load a spatial dataset and create a rgeoda instance
3. how to access attribute data in rgeoowda instance
4. how to create a Queen contiguity weights
5. how to compute local indicators of spatial autocorrelation (LISA) using the Queen contiguity weights
6. how to apply a spatially constrained clustering algorithm (SKATER) to find clusters

For people who using R to do spatial data analysis, a common routine is to use `rgdal` package to read spatial dataset, and then use `sp` package, or (in recent) `sf` package, to handle and analyze spatial data and plot the results on maps.

`rgeoda` provides utility functions to create a rgeoda instance from `sp` object or `sf` object, so that users can easily integrate the functions of spatial data analysis in libgeoda into their existing work.

In this note, I will demonstrate how to work with `sp` package and `libgeoda` package to do spatial data analysis.

## 1. Start from `sp` package

The sp package has been an essential tool which provides spatial data-structures and many utility functions to do spatial analysis in R. It has been a core dependent library for many other packages, e.g. rgdal (IO), maptools (mapping), spdep (spatial weights, spatial statistics, and spatial models) etc. 

Using rgdal to read a ESRI Shapefile will return a `sp` (Spatial object) object, which could be either a SpatialPointsDataFrame (using an AttributeList for its data slot directly), a SpatialLinesDataFrame, or a SpatialPolygonsDataFrame.

```{r}
library(libgeoda)
guerry_path <- system.file("extdata", "Guerry.shp", package = "libgeoda")

library(rgdal)
guerry_sp <- readOGR(guerry_path)
```

You can simply call plot() function to plot this map:

```{r}
plot(guerry_sp)
```

## 2. Create libgeoda object from `sp`

Everything above should be very familiar with you, right? Then, the fun part of spatial data analysis is going to start:

Create a libgeoda object from the `sp` object `guerry_sp` you just created.
```{r message=FALSE}
guerry_gda <- sp_to_geoda(guerry_sp)
```

That's it! It should be very fast: libgeoda uses wkb, which is a binary representation of geometries, to exchange data between `sp` and `libgeoda` in memory. 

## ESDA with libgeoda

Now, with the libgeoda object `guerry_gda`, you can call GeoDa's spatial analysis functions easily. For example, we want to examine the local Moran of variable "crm_prs" (Population per Crime against persons):

```{r}
queen_w <- guerry_gda$CreateContiguityWeights()
crm_lisa <- guerry_gda$LISA(queen_w, as.numeric(paste(guerry_sp$Crm_prs)))
crm_lisa_clusters <- crm_lisa$GetClusterIndicators()
```

Now, with the LISA results, we can do exploratory spatial data analysis by generating a LISA cluster map:

```{r}
lisa_color <- c("#eeeeee","#FF0000","#0000FF","#a7adf9", "#f4ada8", "#999999")
plot(guerry_sp, col=sapply(crm_lisa_clusters, function(x){return(lisa_color[[x+1]])}))
title(main = "Local Moran Map")
legend('bottomright', legend = c("Not Sig", "High-High", "Low-Low", "Low-High","High-Low"), fill = lisa_color)
```
From the above code, you can see that we still use `sp` object to do plotting. The values of cluster indicators from `libgeoda`'s `LISA` object are used to make the LISA map.This emphasis that libgeoda is a feature focused library, which only provides the core functions of spatial data analysis for easy integration.

You can easily append the lisa results to original `sp` object by manipulating the data.frame object inside the `sp` object.

If you check the values of the cluster indicators, you will see they are integer numbers 0 (not significant), 1 (high-high cluster), 2 (low-low cluster), 3 (low-high cluster), 4 (high-low cluster), 5 (neighborless/island), 6 (undefined), which are excatly the same with GeoDa software when you save LISA results to a table:

```{r}
crm_lisa_clusters
```

We also specify what color is associated with which cluster value by using following code:
```
lisa_color <- c("#eeeeee","#FF0000","#0000FF","#a7adf9", "#f4ada8", "#999999")
```
We then create an array of hex color, each observation is set to a hex color according to its cluster indicator value.
```{r}
sapply(crm_lisa_clusters, function(x){return(lisa_color[[x+1]])})
```

